### 2018-11-21 writing/thesis/chapters/methods.tex
Further work on ADER-DG, NS, AMR

% TODO: Extract to settings.tex
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
%\newcommand{\Qv}{\ensuremath{\Qrho^{-1} \Qj}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\potT}{\theta}
\newcommand{\backgroundPotT}{\overline{\theta}}
\newcommand{\pertubationPotT}{\theta'}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}{\vert \lambda_c^{\text{max}} \vert}
\newcommand{\maxViscEigen}{\vert \lambda_v^{\text{max}} \vert}

\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1]{\Phi_{#1}}
\newcommand{\testfunction}[1]{\Phi_{#1}}

@@

\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive at the so called weak formulation of the \textsc{pde} 
\newcommand{\intdt}[1]{\int_{t^n}^{t^{n+1}} #1 \dd{t}}
\newcommand{\intdcell}[1]{\int_{\cell} #1 \dd{\bm{x}}}
\newcommand{\intdcellb}[1]{\int_{\partial{} \cell} #1 \dd{S}} % TODO: define boundary of cell
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\bm{Q}}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(Q, \gradient{Q})} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(Q, \bm{x}, t)
}}
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
\newcommand{\stpredictor}[1]{\bm{q}_{#1}}
We now replace the solution $\bm{Q}$ with so called spacetime-predictor $\stpredictor{h}(\bm{x},t)$ into the weak form and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\intdcell{
  \testfunction{k} \sbasis{l}
}
\right)
\left(\intdt{\intdcellb{
\text{dofs boundary time}
}}\right)
&+\\
\left(\intdt{\intdcell{
\text{Riemann}
}}\right)
-
\left(\intdt{\intdcell{
\text{Flux}
}}\right)
&=
\left(\intdt{\intdcell{
  \text{Source}
}}\right)
\end{split}
\end{align}
The first term motivates an orthogonal basis as this leads to a diagonal mass matrix.

@@

\section{Compressible Navier Stokes}
Following~\cite{dumbser2010arbitrary} by \citeauthor{dumbser2010arbitrary}.

Vector of conserved quantities:
\begin{equation}
  \label{eq:conserved-variables}
 Q = \left( \Qrho, \Qj, \QE \right) 
\end{equation}

The full compressible Navier Stokes equations with heat transfer can be described by
\begin{equation}
  F(Q, \nabla Q) = 
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure + \stressT (Q, \nabla Q)  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure + \stressT (Q, \nabla Q)) - \kappa \nabla T
  \end{pmatrix},
\end{equation}
where $\pressure$, $\stressT$ and $(\kappa \nabla T)$ denote the pressure, stress tensor and heat flux respectively.
Temperature is denoted by $T$.

We use the equation of state of an ideal gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - 0.5 \left(\Qv \cdot \Qj \right) \right).
\end{equation}

The temperature $T$ relates to pressure and density by the ideal gas law
\begin{equation}
  \label{eq:temperature}
 \frac{\pressure}{\Qrho} = RT,
\end{equation}
where $R$ is the specific gas constant.

\todo{Write more text about constants, cite idolikecfd or sth}
\begin{align}
  \begin{split}
  c_v &= \frac{1}{\gamma - 1} R \\
  c_p &= \frac{\gamma}{\gamma - 1} R\\
  R &= c_p - c_v\\
  \gamma &= \frac{c_p}{c_v}
  \end{split}
\end{align}

Heat conduction coefficient $\kappa$
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R
\end{equation}
with $R$ gas constant and $\Pr$ Prandtl number

@@

@@

\section{Diffusive Flux stuff}
Max eigenvalue of convective part $\maxConvEigen$,i.e.\ of $\left( \partial \bm{F}/\partial \bm{Q}\right) \cdot \bm{n}$,
and viscous part $maxViscEigen$, i.e.\ of $\left( \partial \bm{F}/\partial \left( \nabla \bm{Q} \cdot \bm{n} \right)\right) \cdot \bm{n}$
is
\begin{align}
  \maxConvEigen \vert  &= \Vert \Qv \Vert + c\\
  \maxViscEigen \vert &= \max \left( \frac{4}{3} \frac{\mu}{\Qrho}
                                       , \frac{\gamma \mu}{\Pr \Qrho} \right)
\end{align}
with speed of sound $c = \sqrt{\gamma R T }$

Maximum timestep (\textsc{cfl})
\begin{equation}
 \Delta t = \frac{\text{CFL}}{2N + 1} \, \frac{h}{\maxConvEigen + 2 \maxViscEigen \frac{2N+1}{h}}
\end{equation}
with $N$ polynomial order and $h$ characteristic length scale of elements

Numerical flux:
\begin{equation}
  \label{eq:rusanov-flux}
  G(q_h^-, \nabla q_h^-; g_h^+, \nabla q_h^+) \cdot \bm{n} =
  \frac{1}{2} \left(
    F(q_h^+, \nabla q_h^+) +
    F(q_h^-, \nabla q_h^-)
  \right) -
  \frac{1}{2} s_\text{max} (q_h^+ - q_h^-)
\end{equation}
with
\begin{equation}
  \label{eq:parabolic-penalty}
  s_\text{max}  = \max \left(
\vert \lambda_c(q_h^-) \vert, \, \vert \lambda_c(q_h^+)
\right) +
2 \eta \max \left(
\vert \lambda_v(q_h^-) \vert, \, \vert \lambda_v(q_h^+)
\right)
\end{equation}
and
\begin{equation}
  \eta = \frac{N+1}{h}
\end{equation}

@@

We derive a source term by inserting this solution into the \textsc{pde} by using the symbolic math toolkit \textit{SymPy}.
This procedure is called method of manufactured solutions~\cite{salari2000code}.

>>>>

% TODO: Extract to settings.tex
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
%\newcommand{\Qv}{\ensuremath{\Qrho^{-1} \Qj}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\potT}{\theta}
\newcommand{\backgroundPotT}{\overline{\theta}}
\newcommand{\pertubationPotT}{\theta'}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}{\vert \lambda_c^{\text{max}} \vert}
\newcommand{\maxViscEigen}{\vert \lambda_v^{\text{max}} \vert}
\newcommand{\Riemann}{\operatorname{Riemann}}

\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1]{\Phi_{#1}}
\newcommand{\testfunction}[1]{\Phi_{#1}}
\newcommand{\normal}{\bm{n}}
\newcommand{\source}{\bm{S}}

@@

\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive at the so called weak formulation of the \textsc{pde} 
\newcommand{\intdt}[1]{\int_{t^n}^{t^{n+1}} #1 \dd{t}}
\newcommand{\intdcell}[1]{\int_{\cell} #1 \dd{\bm{x}}}
\newcommand{\intdcellb}[1]{\int_{\partial{} \cell} #1 \dd{S}} % TODO: define boundary of cell
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\bm{Q}}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(Q, \gradient{Q})} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(Q, \bm{x}, t)
}}
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
\newcommand{\stpredictor}[1]{\bm{q}_{#1}}
We now replace the solution $\bm{Q}$ with so called spacetime-predictor $\stpredictor{h}(\bm{x},t)$ into the weak form and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\intdcell{
  \testfunction{k} \sbasis{l}
}
\right)
(\bm{u^{n+1} - u^{n}})
+
\left(\intdt{\intdcell{
      \testfunction{k} \Riemann(\bm{q}^-, \bm{q}^+) \cdot \normal
}}\right)
&-\\
\left(\intdt{\intdcell{
    \gradient{\testfunction{k}} \cdot  F(\bm{q})
}}\right)
&=
\left(\intdt{\intdcell{
      \testfunction{k} \source(\bm{q})
}}\right)
\end{split}
\end{align}
The first term motivates an orthogonal basis as this leads to a diagonal mass matrix.

@@

\sidetitle{Diffusive Flux stuff}
Numerical flux:
\begin{equation}
  \label{eq:rusanov-flux}
  \Riemann(q_h^-, \nabla q_h^-; g_h^+, \nabla q_h^+) \cdot \bm{n} =
  \frac{1}{2} \left(
    F(q_h^+, \nabla q_h^+) +
    F(q_h^-, \nabla q_h^-)
  \right) -
  \frac{1}{2} s_\text{max} (q_h^+ - q_h^-)
\end{equation}
with
\begin{equation}
  \label{eq:parabolic-penalty}
  s_\text{max}  = \max \left(
\vert \lambda_c(q_h^-) \vert, \, \vert \lambda_c(q_h^+)
\right) +
2 \eta \max \left(
\vert \lambda_v(q_h^-) \vert, \, \vert \lambda_v(q_h^+)
\right)
\end{equation}
and
\begin{equation}
  \eta = \frac{2N+1}{h \sqrt{\frac{1}{2} \pi}}
\end{equation}

Maximum timestep (\textsc{cfl})
\begin{equation}
 \Delta t \leq  \text{CFL} \, \frac{\alpha(N) \, h}{\maxConvEigen + 2 \maxViscEigen \frac{2N+1}{h}}
\end{equation}
with $N$ polynomial order and $h$ characteristic length scale of elements~\cite{dumbser2010arbitrary,gassner2008discontinuous}.
The constant $\alpha(N) \leq \frac{1}{2N+1}$ is obtained from von Neumann analysis on a simple model problem and depends on the approximation order~\cite{dumbser2008unified}.

\section{Compressible Navier Stokes}
Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary}.

The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 Q = \left( \Qrho, \Qj, \QE \right).
\end{equation}
The flux by
\begin{equation}
  F(Q, \nabla Q) = 
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure + \stressT (Q, \nabla Q)  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure + \stressT (Q, \nabla Q)) - \kappa \nabla T
  \end{pmatrix},
\end{equation}
where $\pressure$, $\stressT$ and $(\kappa \nabla T)$ denote the pressure, stress tensor and heat flux respectively.
Temperature is denoted by $T$.

We close the system with the equation of state of an ideal gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - 0.5 \left(\Qv \cdot \Qj \right) \right).
\end{equation}
The temperature $T$ relates to pressure and density by the ideal gas law
\begin{equation}
  \label{eq:temperature}
 \frac{\pressure}{\Qrho} = RT,
\end{equation}
where $R$ is the specific gas constant.

We further introduce the ratio of specific heats $\gamma$ and the heat fraction for constant volume $c_v$ and constant volume $c_r$.
These constants are all fluid dependent and relate to each other by
\begin{align}
  \begin{split}
  c_v &= \frac{1}{\gamma - 1} R \\
  c_p &= \frac{\gamma}{\gamma - 1} R\\
  R &= c_p - c_v\\
  \gamma &= \frac{c_p}{c_v}.
  \end{split}
\end{align}
\todo{Cite sth!}

We can finally define the heat conduction coefficient $\kappa$
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R
\end{equation}
where Prandtl number $\text{Pr}$ again depends on the fluid.

@@

Max eigenvalue of convective part $\maxConvEigen$,i.e.\ of $\left( \partial \bm{F}/\partial \bm{Q}\right) \cdot \bm{n}$,
and viscous part $maxViscEigen$, i.e.\ of $\left( \partial \bm{F}/\partial \left( \nabla \bm{Q} \cdot \bm{n} \right)\right) \cdot \bm{n}$
is
\begin{align}
  \maxConvEigen \vert  &= \Vert \Qv \Vert + c\\
  \maxViscEigen \vert &= \max \left( \frac{4}{3} \frac{\mu}{\Qrho}
                                       , \frac{\gamma \mu}{\Pr \Qrho} \right)
\end{align}
with speed of sound $c = \sqrt{\gamma R T }$

@@

\section{Adaptive Mesh Refinement}
\label{sec:amr}
We use adaptive mesh refinement (\textsc{amr}).

Use total variation (\textsc{tv}) given by
\begin{equation}
  \label{eq:tv}
  \operatorname{TV}\left( f \right) =
\intdcell{ \vert \gradient{f} \vert }
\end{equation}

We introduce parallel reduce.

Reduce mean/std by~\cite{chan1982updating}

Criterion with threshold.
Justify by Chebychev's inequality?

@@

We derive a source term by inserting this solution into the \textsc{pde} by using the symbolic math toolkit \textit{Sage}.
This procedure is called method of manufactured solutions~\cite{salari2000code}.

### 2018-11-21 writing/thesis/pages/abstract.tex
Setup latex template.
Further work on ADER-DG, NS, AMR

\chapter{Methods}
Methods here.

\todo{Organize into chapters!}

>>>>

\chapter{\abstractname}
This thesis is concerned with the simulation of clouds with the ExaHyPE framework.
ExaHyPE is an engine for hyperbolic PDEs that uses the ADER-DG scheme.

We use the compressible Navier-Stokes (NS) equations with heat conduction to describe the dynamics of clouds.
Due to the diffusive components, the NS-Equations are not hyperbolic.
We thus need to modify parts of the numerical scheme.
